% ===============================
% Document Class and Options
% ===============================
% \documentclass[14pt, letterpaper, twoside, openright, draft]{memoir}
\documentclass[11pt, statementpaper, twoside, openright]{memoir}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
% \usepackage{lmodern}
\usepackage{Baskervaldx}
\usepackage{caption}
\captionsetup[figure]{labelformat=empty, font={footnotesize, it}, justification=centering}

% Remove chapter numbering from headings and TOC
\makeatletter
\renewcommand{\@chapapp}{}           % no "Chapter" label
\renewcommand{\thechapter}{}         % no chapter number
\makeatother

% Define a custom chapter style that omits numbering
\makechapterstyle{noNumber}{
  \chapterstyle{default}
  \renewcommand{\printchaptername}{}
  \renewcommand{\printchapternum}{}
  \renewcommand{\afterchapternum}{}
}
\chapterstyle{noNumber}

% Remove the period before the chapter name in headers
\makeatletter
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\makeatother

% Adjust header style
\pagestyle{headings} % Ensure that the document uses the "headings" style

% Adjust section numbering depth so chapters aren’t numbered in TOC
\setcounter{secnumdepth}{-1}
\setcounter{tocdepth}{0}  % chapters appear in TOC without numbering

% --- Page Layout and Geometry ---
\setlrmarginsandblock{1in}{.75in}{*}         % Set left/right margins
\setbinding{0.25in}                            % Add gutter for binding
\setulmarginsandblock{1in}{1in}{*} % Set top and bottom margins to 1 inch
\checkandfixthelayout   
\linespread{1.2}                                 % Set line spacing

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{csquotes}

\title{Pride and Prejudice}
\author{Jane Austen}
\date{} % Leave the date blank to suppress it

\begin{document}
\frontmatter
\maketitle
\thispagestyle{empty} % Ensures no headers or footers
\include{copyright.tex}
\include{preface.tex}
\tableofcontents

\chapter{List of Illustrations}
PAGE

Frontispiece iv

Title-page v

Dedication vii

Heading to Preface ix

Heading to List of Illustrations xxv

Heading to Chapter I. 1

``He came down to see the place'' 2

Mr. and Mrs. Bennet 5

``I hope Mr. Bingley will like it'' 6

``I'm the tallest'' 9

``He rode a black horse'' 10

``When the party entered'' 12

``She is tolerable'' 15

Heading to Chapter IV. 18

Heading to Chapter V. 22

``Without once opening his lips'' 24

Tailpiece to Chapter V. 26

Heading to Chapter VI. 27

``The entreaties of several'' 31

``A note for Miss Bennet'' 36

``Cheerful prognostics'' 40

``The apothecary came'' 43

``Covering a screen'' 45

``Mrs. Bennet and her two youngest girls'' 53

Heading to Chapter X. 60

``No, no; stay where you are'' 67

``Piling up the fire'' 69

Heading to Chapter XII. 75

Heading to Chapter XIII. 78

Heading to Chapter XIV. 84

``Protested that he never read novels'' 87

Heading to Chapter XV. 89

Heading to Chapter XVI. 95

``The officers of the ------shire'' 97

``Delighted to see their dear friend again'' 108

Heading to Chapter XVIII. 113

``Such very superior dancing is not often seen'' 118

``To assure you in the most animated language'' 132

Heading to Chapter XX. 139

``They entered the breakfast-room'' 143

Heading to Chapter XXI. 146

``Walked back with them'' 148

Heading to Chapter XXII. 154

``So much love and eloquence'' 156

``Protested he must be entirely mistaken'' 161

``Whenever she spoke in a low voice'' 166

Heading to Chapter XXIV. 168

Heading to Chapter XXV. 175

``Offended two or three young ladies'' 177

``Will you come and see me?'' 181

``On the stairs'' 189

``At the door'' 194

``In conversation with the ladies'' 198

``Lady Catherine,'' said she, ``you have given me a treasure'' 200

Heading to Chapter XXX. 209

``He never failed to inform them'' 211

``The gentlemen accompanied him'' 213

Heading to Chapter XXXI. 215

Heading to Chapter XXXII. 221

``Accompanied by their aunt'' 225

``On looking up'' 228

Heading to Chapter XXXIV. 235

``Hearing herself called'' 243

Heading to Chapter XXXVI. 253

``Meeting accidentally in town'' 256

``His parting obeisance'' 261

``Dawson'' 263

``The elevation of his feelings'' 267

``They had forgotten to leave any message'' 270

``How nicely we are crammed in!'' 272

Heading to Chapter XL. 278

``I am determined never to speak of it again'' 283

``When Colonel Miller's regiment went away'' 285

``Tenderly flirting'' 290

The arrival of the Gardiners 294

``Conjecturing as to the date'' 301

Heading to Chapter XLIV. 318

``To make herself agreeable to all'' 321

``Engaged by the river'' 327

Heading to Chapter XLVI. 334

``I have not an instant to lose'' 339

``The first pleasing earnest of their welcome'' 345

The Post 359

``To whom I have related the affair'' 363

Heading to Chapter XLIX. 368

``But perhaps you would like to read it'' 370

``The spiteful old ladies'' 377

``With an affectionate smile'' 385

``I am sure she did not listen'' 393

``Mr. Darcy with him'' 404

``Jane happened to look round'' 415

``Mrs. Long and her nieces'' 420

``Lizzy, my dear, I want to speak to you'' 422

Heading to Chapter LVI. 431

``After a short survey'' 434

``But now it comes out'' 442

``The efforts of his aunt'' 448

``Unable to utter a syllable'' 457

``The obsequious civility'' 466

Heading to Chapter LXI. 472

The End 476


\mainmatter
\include{main_matter.tex}
\end{document}